\documentclass{article}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{fullpage}
\usepackage[shortlabels]{enumitem}


\makeatletter
\def\moverlay{\mathpalette\mov@rlay}
\def\mov@rlay#1#2{\leavevmode\vtop{%
   \baselineskip\z@skip \lineskiplimit-\maxdimen
   \ialign{\hfil$\m@th#1##$\hfil\cr#2\crcr}}}
\newcommand{\charfusion}[3][\mathord]{
    #1{\ifx#1\mathop\vphantom{#2}\fi
        \mathpalette\mov@rlay{#2\cr#3}
      }
    \ifx#1\mathop\expandafter\displaylimits\fi}
\makeatother

\newcommand{\cupdot}{\charfusion[\mathbin]{\cup}{\cdot}}
\newcommand{\bigcupdot}{\charfusion[\mathop]{\bigcup}{\cdot}}


\title{MA 2631 Conference 6}

\setlength\parindent{0pt}
\begin{document}
\maketitle



\begin{enumerate}

%%% 1 %%%
\item

In a small town, there are 50 births a year. Assume that the probability that a newborn is a girl is 50\%. How likely is it that in a given year, there are at least 25 and at most 27 girls born. Calculate this probability
\begin{itemize}
	\item[a)] exactly.
	\item[b)] by an approximation with the normal distribution.
\end{itemize}


%%% 2 %%%
\item
Consider a biased coin that shows {\em heads} in $\frac{2}{3}$ of all cases and {\em tails} only in $\frac{1}{3}$ of all cases. The coin is flipped consecutively (and independently) 200 times. 
\begin{itemize}
	\item[a)] What is the probability that {\em tails} shows up the first time at the 10th flip?
	\item[b)] Calculate the probability  {\em heads}  shows up more than 150 times (using a suitable approximation). 

\end{itemize}

%%% 3 %%%
\item 

Assume that the joint probability mass distribution $p_{X,Y}$ of the random variable $X$ and $Y$ is given by
\begin{align*}
p_{X,Y}(0,0) =& \frac{1}{3} \qquad p_{X,Y}(0,1)= \frac{1}{4};\\
p_{X,Y}(1,0) =& \frac{1}{4} \qquad  p_{X,Y}(1,1) = \frac{1}{6}.
\end{align*}

\begin{itemize}
	\item[a)] Calculate the marginal probability mass distributions $p_X$ and $p_Y$.
%	\item[b)] Calculate the covariance and correlation of $X$ and $Y$
	\item[b)] What is the probability mass distribution of the random variable $Z = X^2 + Y$?
\end{itemize}

%%% 4 %%%

\item

Assume that $X$ and $Y$ are jointly distributed random variables with joint density 
\[
f_{X,Y}(x,y) = \left\{ \begin{array}{ll} cy e^{-x} & \mbox{if } 0 \leq x < \infty, \, 0 \leq y \leq 1;\\ 0  & \mbox{else}.\end{array} \right.
\]
for some $c \in \mathbb{R}$.
\begin{itemize}
	\item[a)] Calculate $c$.
	\item[b)] Calculate the marginal probability density functions $f_X$ and $f_Y$.
	\item[c)] What is the probability $P[ 3X+ Y^2 > 4]$?
\end{itemize}

\end{enumerate}

\end{document}